---
title: "HW 8"
author: "Marley Akonnor"
date: "9/6/2021"
output: pdf_document
---

Step 1: Load the Data
Access read.xls function from the gdata package
```{r}
library(gdata)
```

Read the data
```{r}
antelope_pop <- read.xls("/Users/m/Documents/M.S Syracuse Data Science/Courses/IST 687 - Intro to Data Science/Week 9/mlr01.xls")
```

Rename the columns
```{r}
new_column_names <- c("Fawn_Population", "Antelope_Population", "Annual_Precipitation", "Winter_Severity")
```

Assign the column names to the dataset
```{r}
colnames(antelope_pop) <- new_column_names
```

Inspect the dataset
```{r antelope_pop}
str(antelope_pop)
```

Create bivariate plots of number of baby fawns versus adult antelope population, the precipitation that year, and the severity of the winter.
fawn(y) vs antelope(x)
```{r Fawn Population Growth 1}
plot(antelope_pop$Antelope_Population,antelope_pop$Fawn_Population)
```

fawn(y) vs precipitation(x)
```{r Fawn Population Growth 2}
plot(antelope_pop$Annual_Precipitation,antelope_pop$Fawn_Population)
```

fawn(y) vs winter(x)
```{r Fawn Population Growth 3}
plot(antelope_pop$Winter_Severity,antelope_pop$Fawn_Population)
```

Next, create three regression models of increasing complexity using lm()
In the first model, predict the number of fawns from the severity of the winter
```{r}
fawn_wint_model_1 <- lm(formula = Fawn_Population ~ Winter_Severity, data = antelope_pop)
summary(fawn_wint_model_1)
```

In the second model, predict the number of fawns from two variables (one should be the severity of the winter)
```{r}
fawn_precip_wint_model_2 <- lm(formula = Fawn_Population ~ Annual_Precipitation + Winter_Severity, data = antelope_pop)
summary(fawn_precip_wint_model_2)
```

In the third model predict the number of fawns from the three other variables
```{r}
all_var_model_3 <- lm(formula = Fawn_Population ~ Antelope_Population + Annual_Precipitation + Winter_Severity, data = antelope_pop)
summary(all_var_model_3)
```

Which model works best?
The third model works the best. The multiple R-squared is 97% and the F statistic P-value is the lowest of the 3.

Which of the predictors are statistically significant in each model?
The statistically significant predictor in model 1 is: Winter Severity
The statistically significant predictor in model 2 is: Annual Precipitation
The statistically significant predictors in model 3 are: Winter Severity, Annual Precipitation, and Antelope Population

If you wanted to create the most parsimonious model (i.e., the one that did the best job with the fewest predictors), what would it contain?
It would contain the Antelope Population and Annual Precipitation.

