---
title: "Olympics Project"
author: "Marley Akonnor"
date: "9/14/2021"
output:
  pdf_document: default
  html_document: default
---

Question: Can we predict the Medal won based on Age, Height, and Weight
```{r}
library(neuralnet)
```
Read in the dataset
```{r}
oly_data <- read.csv("/Users/m/Documents/M.S Syracuse Data Science/Courses/IST 687 - Intro to Data Science/Project/athlete_events.csv")
```

#Clean the dataset - rid of NAs
```{r}
oly_data <- na.omit(oly_data)
```

Check to see if any remain
```{r}
sum(is.na(oly_data))
```

Inspect the structure
```{r}
str(oly_data)
```

Look at a summmary of our data
```{r}
summary(oly_data)
```

Convert the "Sport" column from characters to factors
```{r}
oly_data$Sport <- as.factor(oly_data$Sport)
```

What are the unique values of medal?
```{r}
unique(oly_data$Medal)
```

Make "Medal" column easier to read and convert to int
```{r}
oly_data[oly_data$Medal == "Gold", ]$Medal <- 0
oly_data[oly_data$Medal == "Silver", ]$Medal <- 1
oly_data[oly_data$Medal == "Bronze", ]$Medal <- 2
oly_data$Medal <- as.numeric(oly_data$Medal)
```

Examine distributions with histograms
```{r}
hist(oly_data$Age)

hist(oly_data$Height)

hist(oly_data$Weight)
```

Min-Max normalization
create Normalization function (https://datasharkie.com/how-to-normalize-data-in-r/)
```{r}
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}
```

Normalize Age and look at histogram
```{r}
oly_data$Age <- normalize(oly_data$Age)
hist(oly_data$Age)
```

Normalize Height and look at histogram
```{r}
oly_data$Height <- normalize(oly_data$Height)
hist(oly_data$Height)
```

Normalize Weight and look at histogram
```{r}
oly_data$Weight <- normalize(oly_data$Weight)
hist(oly_data$Weight)
```

Normalize Medal and look at histogram
```{r}
oly_data$Medal <- normalize(oly_data$Medal)
hist(oly_data$Medal)
```

Create a new Olympic dataframe with relevant variables
```{r}
new_oly_data <- data.frame(oly_data$Medal, oly_data$Age, oly_data$Height, oly_data$Weight)
```

Rename the columns
```{r}
col_names <- c("Medal", "Age", "Height", "Weight")
colnames(new_oly_data) <- col_names
```

Data partition and random indexing
```{r}
ran_ind <- sample(1:dim(new_oly_data)[1])
prorat_data <- floor(2 * dim(new_oly_data)[1]/3)
```

Create training data
```{r}
training_data <- new_oly_data[ran_ind[1:prorat_data], ]
```

Create test data
```{r}
testing_data <- new_oly_data[ran_ind[(prorat_data+1):dim(new_oly_data)[1]],]
```

Neural Network model
```{r}
oly_nn <- neuralnet(Medal ~ Age + Height + Weight, data = training_data, hidden = 2, err.fct = "sse", linear.output = FALSE)
plot(oly_nn, rep = "best")
```

Prediction
```{r}
nn_output <- compute(oly_nn, training_data[, -1])
head(nn_output$net.result)
head(training_data[1:6, ])
```

Create a subset of the testing data
```{r}
temp_test <- subset(testing_data, select = c("Age", "Height", "Weight"))
head(temp_test)
```

Run Neural Net on test data
```{r}
olynn.results <- compute(oly_nn, temp_test)
results <- data.frame(actual=testing_data$Medal, prediction=olynn.results$net.result)
results[1:20,]
```
